Supporting Information : Impact of asymmetric tail-dependence in extreme climatic events on extinction risk in population ecology

### 2019-03-16 MainText.Rmd
suppmat description of density dependence model : upto Hassell added

<!--
Nature of tail dependence in environmental noise (or in vital rates) affects extinction risk of a population differently.
       - Age structure stochastic matrix model : extrisk of left tail dep. > extrisk of right tail dep.
         (This difference is small, could be explained by Tuljapurkar's approx. for linear model, see discussion)

       - 6 Density-dependent models with extreme tail dep. in noise (local/ global dispersal)
           - extrisk depends on tail dependnece as well as on model parameters (under-compensatory, over-compensatory)
           - Rescue effect ??
           - H0 : switching occurs near rc (without dispersal)
-->

We explored the potential of asymmetric tail-dependence in copula 
for understanding the extinction risk of populations towards extreme climatic event in ecology
by addressing the following specific questions: (Q1) Does nature of tail dependence in vital rates
of a single-patch age-structured population model affect extinction risk of that population differently?
We hypothesized population might have higher extinction risk if vital rates are left-tail dependent.
(Q2) How does asymmetric tail-dependent noise across multiple patches influence extinction 
risk of a nonlinear metapopulation model having density dependence? We hypothesized the general finding 
from density-independent metapopulation model (higher extinction risk if noise across patches were 
left-tail dependent than the one if noise came from right-tail dependent) holds true only in
the under-compensatory regime, but should be reverse if the model shows over-compensatory dynamics. 
(Q3) We further hypothesize that when there is no dispersal, the above-said switching should occur
near the critical parameter value where the deterministic model shows transition from 
under-compensatory to over-compensatory dynamics.

>>>>

We explored the potential of asymmetric tail-dependence in copula 
for understanding the extinction risk of populations towards extreme climatic event in ecology
by addressing the following specific questions: (Q1) Does nature of tail dependence in vital rates
of a single-patch age-structured population model affect extinction risk of that population differently?
We hypothesized population might have higher extinction risk if vital rates are left-tail dependent.
(Q2) How does asymmetric tail-dependent noise across multiple patches influence extinction 
risk of a nonlinear metapopulation model having density dependence? We hypothesized the general finding 
from density-independent metapopulation model [BIVAN paper results of Lewinton-Cohen model]
(higher extinction risk if noise across patches were 
left-tail dependent than the one if noise came from right-tail dependent) holds true only in
the under-compensatory regime, but should be reverse if the model shows over-compensatory dynamics. 
(Q3) We further hypothesize that when there is no dispersal, the above-said switching should occur
near the critical parameter value where the deterministic model shows transition from 
under-compensatory to over-compensatory dynamics.

### 2019-03-16 SuppMat.Rmd
suppmat description of density dependence model : upto Hassell added

dt2<-tibble("Density dependence $(\\lambda_t^{i})$ with noise" = c("$\\exp(\\epsilon_t^{i}) \\hspace{0.1 cm} \\exp[r (1-P_t^{i}/K)$]",
                          "$\\exp(\\epsilon_t^{i}) \\hspace{0.1 cm} r/(1+aP_t^{i})^b$",
                          "$\\exp(\\epsilon_t^{i}) \\hspace{0.1 cm} r/[1+(aP_t^{i})^b]$",
                          "$\\exp(\\epsilon_t^{i}) \\hspace{0.1 cm} r/[1+\\exp \\{-a(1-P_t^{i}/b)\\}]$",
                          "$\\exp(\\epsilon_t^{i}) \\hspace{0.1 cm} [1+r(1-P_t^{i})]$",
                          "$\\exp(\\epsilon_t^{i}) \\hspace{0.1 cm} [1+r(K-P_t^{i})\\{1-\\exp(-sP_t^{i})\\}]$"
))

@@

>>>>

dt2<-tibble("Density dependence $(\\lambda_t^{i})$ with noise $(\\epsilon_t^{i})$" = c("$\\exp(\\epsilon_t^{i}) \\hspace{0.1 cm} \\exp[r (1-P_t^{i}/K)$]",
                          "$\\exp(\\epsilon_t^{i}) \\hspace{0.1 cm} r/(1+aP_t^{i})^b$",
                          "$\\exp(\\epsilon_t^{i}) \\hspace{0.1 cm} r/[1+(aP_t^{i})^b]$",
                          "$\\exp(\\epsilon_t^{i}) \\hspace{0.1 cm} r/[1+\\exp \\{-a(1-P_t^{i}/b)\\}]$",
                          "$\\exp(\\epsilon_t^{i}) \\hspace{0.1 cm} [1+r(1-P_t^{i})]$",
                          "$\\exp(\\epsilon_t^{i}) \\hspace{0.1 cm} [1+r(K-P_t^{i})\\{1-\\exp(-sP_t^{i})\\}]$"
))

@@

First, we will briefly discuss about the stability analysis for a difference 
equation which is frequently found in population ecology model. Let's say, $P_{t+1} = F(P_t)$
represents the model dynamics of a single population $P$ which progress recursively via 
density-dependence function $F$ with discrete time variable $t$. Fixed-point analysis can
be used to get information about this type of nonlinear difference equation. The idea is to
employ the concept that the system at fixed point (or equilibrium $P_e$) does not change with time 
(i.e. $P_{t+1} = P_t$). Given a tiny perturbation from its equilibrium $P_e$, if the system 
returns to its original state, then $P_e$ is called to be a stable equilibrium, otherwise unstable.
Mathematically, from standard linear stability analysis we can say an equilibrium is locally 
asymptotically stable if $| {F}' | < 1$, where ${F}' = \left( \frac{\mathrm{d}F }{\mathrm{d} P} \right) _{P_t=P_e}$.
If the approach towards its equilibrium is monotonic (when $0<{F}'<1$), the dynamics is said to be
in under-compensatory regime. An over-compensatory dynamics means the approach towards its stable equlibrium
is oscillatory ($-1<{F}'<0$). Following we will describe stability condition of the model dynamics considered for each 
patch of metapopulation models in the maintext (see Table \ref{MT-tab_model_summary}).

**Ricker model** [@ricker1954stock], originally introduced in the context of modelling
stock and recruitment of salmon populations, is a classic 
discrete population model 
and frequently used in fisheries management problem. The model dynamics
is represented as 
$P_{t+1}=P_t \exp[r(1-P_t/K)]$, where $P$ is the population density, $r>0$ is the intrinsic 
growth rate and $K>0$ is the carrying capacity. There are two equilibrium states ($P_e = 0, K$).
The equilibrium at ($P_e = 0$) is unstable as ${F}' = \exp(r) > 1$ and considered 
as the extinction equilibrium 
in population ecology context. Depending on parameter $r$, 
the other equilibrium $P_e = K$ can be stable or unstable as ${F}' = (1-r)$. Following generalized
concept, the model can show monotonic approach to its stable equilibrium $P_e = K$ if $0<r<1$ 
or can approach $P_e = K$ in an oscillating fashion if $1<r<2$. When 
$r>2$, system shows sustained oscillation and the equilibrium becomes unstable. We called $r=1$ 
as $r_c$, the critical point where system goes from under- to over-compensatory regime and 
$r=2$ as $r_{bf}$ where bifurcation (transition from stable to unstable equilibrium) occurs.

**Hassell model** [@hassell1975density], motivated for modelling insect populations,
can be represented as $P_{t+1} = rP_t/(1+aP_t)^b$, where $r>0$ is the intrinsic growth rate 
of population density $(P)$. $a$ and $b$ are the other positive parameters. If $b=1$, it reduces to
Beverton-Holt model [@beverton2012dynamics] used in fisheries. In general, if $b \sim 1$,
then we say that there is contest and $F$ will saturate eventually at a finite
level for large populations, which indicates that the population stabilizes by
rejecting too many newborns. For scramble, $b$ is large and $F$ tends to
zero for large populations which indicates that the resources are over-utilized
leading to eventual extinction. Like Ricker, Hassel model too has two equilibrim:
$P_e =0$ (trivial extinction equilibrium) and $P_e = P^\star =(r^{1/b}-1)/a$. 
For $P_e = P^\star$, model remains in under-compensatory regime if $1<r<[b/(b-1)]^b$ 
and in over-compensatory regime if $[b/(b-1)]^b<r<[b/(b-2)]^b$ with $r_c = [b/(b-1)]^b$.
Hopf bifurcation occurs at $r_{bf} = [b/(b-2)]^b$, so if $r>r_{bf}$, $P_e=P^\star$
becomes unstable equilibrium.

**Maynard Smith model**

**Pennycuick model**

**Verhulst model**

**Austin-Brewer model**

# References

\clearpage
